---
title: "Brownlows Bets"
author: "Sisilia Martinez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "7/5/2022"    
    
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(tigerstats)
require(knitr)
source(file="BrLw.R")
```


<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 637 times so far. Where do you think his profits are? Discuss this in a markdown document
 
Report git commit structure:
There are 3 questions and there should be at least 3 separate commits, at least one dealing with each of the 3 questions.
1. What are the fair odds for a bet on matching birthdays in a 35 person class.
2. For what size class is a 2:1 odds bet fair.
3. How much money would Brownlow be expected to end up with if the bets went as described.

## Our strategy is to build a Function and the use it to answer the three questions. 

The function will be given a class_size by number of students, odds, and Number_of_times brownlow made the bet.  It will give you back how much Brownlows earnings were. We will use it to answer our three questions below. If you can, improve this function by giving default values to some of its inputs. 

# In the class and payoff as described  If you highered 40 brownlow substitute teachers to do the same thing that Brownlow did then.
## what is their average earnings
## the standard deviation and a 95% confidence interval for their earnings?
  
<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows winnings (or losings)-->

We are creating an array with 40 different substitute teachers doing the same bet Brownlow did, and than we will find the average that the teachers made.

```{r}
Array40 <- vector("numeric")
for(i in 1:40){
  Array40[i]<- BrLw(numClass = 35, odds = 2, countoDo = 637)
}

favstats(Array40)
lcl <- mean(Array40) - 2*sd(Array40)
ucl <- mean(Array40) + 2*sd(Array40)
options(digits = 1)
print(paste(lcl, ucl))
```

The lower end is `r lcl` to the upper end `r ucl`. If Brownlow was to give out a 2:1 bets the lower end he would be making is `r lcl` and the upper end would be `r ucl`. This is not a fair bet because he will constantly keep making money. 

# What are the fair odds? Use manipulation of odds in the function to tune Brownlows earnings to zero. 


<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 637 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->

```{r}
Odds40 <- vector("numeric")
for(i in 1: 100)(
  Odds40[i] <- BrLw(odds = 4.3999)
)

favstats(Odds40)
lcl <- mean(Odds40) - 2*sd(Odds40)
ucl <- mean(Odds40) + 2*sd(Odds40)
options(digits = 1)
print(paste(lcl, ucl))
```

The fair odds for Brownlow's bets are 4.3999. This gave us a close average to 0. By having it close to 0 it allows us to see that these odds are fair because he is not earning anything on average. The mean being `r mean(Odds40)`, the lowest amount being `r lcl`, and the highest amount made being `r ucl`.

# What size of class makes the 2:1 bet fair?  Use manipulation of number of Students to tune Brownlows earnings to zero.

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting profits!-->


```{r}
Class40 <- vector("numeric")
for(i in 1: 100)(
  Class40[i] <- BrLw(numClass = 28)
)

favstats(Class40)
lcl <- mean(Class40) - 2*sd(Class40)
ucl <- mean(Class40) + 2*sd(Class40)
options(digits = 1)
print(paste(lcl, ucl))
```

In order to make the bet fair the size of class should be 28. This gives us an average close to 0 with notifying the class size. With the mean being `r mean(Class40)`, the lowest amount made `r lcl`, and the highest amount made `r ucl`.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

Since Brownlow's original 2:1 bet had him constantly making money, it is ruled out as a not fair bet. In order to make this bet fair we would have to set the odds to 4.3999 to make the average closer to 0. Just like the odds in order to make this bet fair we will also have change the size of the class from 30 to 28. We change it to 28 to make the average closer to 0 so that he will not constantly make money. The lowest amount he will make with these modifications would be `r lcl` and the highest amount he will make would be `r ucl`.

